---
title: Docker AWS
author: Mike Gahan
date: '2017-11-02'
slug: docker-aws
categories:
  - R
tags:
  - R
  - AWS
  - Docker
---

```{r setup, include=FALSE}
library(knitr)
library(data.table)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE)
```

## If you would like to follow along:

This tutorial assumes the user has already downloaded and installed the following:

- Docker
	- Mac OSX: https://docs.docker.com/docker-for-mac/
	- Windows: https://docs.docker.com/engine/installation/windows/
- AWS Command Line Interface (CLI)
	- https://aws.amazon.com/cli/

## Why docker is useful?

- No worries about missing dependencies, packages, and other pain points during
subsequent deployments

- Reduce/eliminate concerns about compatibility on different platforms. This has the potential
to eliminate much of our Windows/Mac issues.

- Allows access to a VM without all the overhead of a VM

- Allows us to quickly scale an environment through AWS or other cloud platform
quickly and without pain

## Creating docker machine

A nice feature to docker is that is can be run to setup on new environment on your
local machine, or it can also be used to create a new machine via the AWS driver
(or any other of the available drivers such as Google, Microsoft Azure, Digital Ocean,
Rackspace, etc.).

Here we will begin by creating a machine using AWS.
AWS has a number of instance-type options for a variety of situations. As of October 26, 2016,
here are the instance types and their respective price per hour.

```{r, echo=FALSE, eval=FALSE}
instanceTypes <- fread("instanceTypes.csv", showProgress = FALSE)
kable(instanceTypes)
```

A more up to date view of this can be at the link below:

https://aws.amazon.com/ec2/pricing/on-demand/

Now we can build a machine using docker. Key parameters we need to input:

- **Access Key** - security access key. Ask admins for this if not available.
- **Secret Key** - security secret key. Again, ask admins for this if not available.
- **Region** - default aws region associated with access and secret key
- **Root Size** - size in gigabytes of drive space. For RStudio instances, it is nice
to have 15-20 GBs and more if you are processing lots of data
- **Instance Type** - type of instance desired
- **Security Group** - optional input. For the modeling team, the docker-machine-ips
only allows access from certain IP addresses. Here is an example of how to open we might
want to open ports:

```
The following ports must be open in the AWS security group
Type	        Protocol	Port Range		Source
HTTP	        TCP  	        80        0.0.0.0/0  
SSH             TCP             22        0.0.0.0/0
Custom TCP Rule TCP           2376        0.0.0.0/0
Custom TCP Rule TCP           8787        0.0.0.0/0
Custom TCP Rule TCP           3838        0.0.0.0/0
```

There is more detail on these parameters at the following link:
https://docs.docker.com/machine/drivers/aws/

Creating the machine can take a couple of minutes to build the environment, certificates,
and SSH keys.

```{sh, eval=FALSE}
name=Mike
echo $name

docker-machine create --driver amazonec2 \
--amazonec2-access-key A******************** \
--amazonec2-secret-key ********************* \
--amazonec2-region us-west-2 \
--amazonec2-root-size 20 \
--amazonec2-instance-type t2.micro \
 awsRStudio$name
```

Note: You can also provide a security group to the following command (e.g):

```{sh, eval=FALSE}
--amazonec2-security-group docker-machine-ips \
```

```
Running pre-create checks...
Creating machine...
(awsRStudio) Launching instance...
Waiting for machine to be running, this may take a few minutes...
Detecting operating system of created instance...
Waiting for SSH to be available...
Detecting the provisioner...
Provisioning with ubuntu(systemd)...
Installing Docker...
Copying certs to the local machine directory...
Copying certs to the remote machine...
Setting Docker configuration on the remote daemon...
Checking connection to Docker...
Docker is up and running!
To see how to connect your Docker Client to the Docker Engine running on this virtual machine, run: docker-machine env awsRStudio
```

## Activate new docker machine

Now we have created a new machine, but this does not mean that we this is the current
active machine. Let's list the docker machines available to us.

```{sh, eval=FALSE}
docker-machine ls
```

```
NAME         ACTIVE   DRIVER       STATE     URL                        SWARM   DOCKER    ERRORS
awsRStudio   -        amazonec2    Running   tcp://35.161.235.82:2376           v1.12.2   
default      -        virtualbox   Stopped                                      Unknown  
```

The `ACTIVE` column puts a `*` next to the active machine. The new machine we have created
is not currently the active machine. We should make it the active machine.

```{sh, eval=FALSE}
eval $(docker-machine env awsRStudio$name)
```

So now we have an activated machine! We should soon be able to run our code but we will
need an image to work off first.

```
NAME         ACTIVE   DRIVER       STATE     URL                        SWARM   DOCKER    ERRORS
awsRStudio   *        amazonec2    Running   tcp://35.161.235.82:2376           v1.12.2   
default      -        virtualbox   Stopped                                      Unknown  
```


## Pull Image from Repository

In order to eventually run code in our docker containers, we need to pull or create
an image that will eventually spawn a container. There are a couple of different images
we can pull.

1. Modeling team RStudio/Shiny Image
	- this image comes complete with a RStudio-server and a Shiny server as well as all
	AWS profiles and individual modeling team member profiles
	
2. Generic docker hub RStudio image
	- comes with RStudio-server and common R packages
	- only comes with rstudio profile (username: rstudio, password: rstudio)
	

In order to grab the first image, try the following code. This code is used
to login to the AWS container:

```{sh, eval=FALSE}
eval $(aws ecr --profile metabiota_internal get-login --region us-west-2)
```

```
Flag --email has been deprecated, will be removed in 1.13.
Login Succeeded
```

If that didn't work, that is okay. We can get you setup later. For the time being we will
grab the generic RStudio-Server image from docker hub.

1. Pull Modeling team RStudio/Shiny Image

```{sh, eval=FALSE}
docker pull 277815483075.dkr.ecr.us-west-2.amazonaws.com/metabiota/rstudio:latest
```

2. or Pull Generic RStudio-Server image

```{sh, eval=FALSE}
docker pull rocker/rstudio
```

The image can take anywhere from a few seconds to a few minutes to download and then extract
sections of the image.

## Create container from image

In docker lingo, an image does not actually run code, rather containers do. Images
are immutable meaning they can't be altered unless they are rebuilt. We have
already downloaded an image but we have not yet built a container. We can
build containers from an image using `docker run`.

Buiding the container depends on the image you pulled in the previous step.

1. Build container from Modeling team RStudio/Shiny Image

```{sh, eval=FALSE}
docker run -d -p 8787:8787 -p 3838:3838 --name rstudio_test --shm-size=1g 277815483075.dkr.ecr.us-west-2.amazonaws.com/metabiota/rstudio:latest
```

2. or Build container from Generic RStudio-Server Image

```{sh, eval=FALSE}
docker run -d -p 8787:8787 --name rstudio_test --shm-size=1g rocker/rstudio:latest
```

There is a lot here! We can go ahead each flag independently.

- **-d detached** This runs the container in detached mode (in the background). 
Very desireable for apps like RStudio-Server or Shiny-Server that run through ports that
can be access through a browser.

- **-p port** What ports should be opened and where should they be mapped? RStudio-Server
defaults to being mapped through port 8787. If we wanted to view RStudio through port
8788, we would change this to 8787:8788.

- **--name name** What is the name of the container? If container is not named,
docker assigns a name based on a random adjective and scientist (e.g. *drunk_einstein*)

- **--shm-size shared memory** I think this is especially an important flag for
the modeling team since we read files into R using the aws cli. That process is
technically sharing memory from R to the system calls. I usually max this out at the amount
of memory available.

## Naviage to RStudio-Server

In order to actually navigate to our newly built RStudio-Server, we need to find the
IP where the container is running. First, let's make sure our container is up and running.

```{sh, eval=FALSE}
docker ps
```

```
CONTAINER ID        IMAGE                                                                   COMMAND             CREATED             STATUS              PORTS                                            NAMES
f00fa3ff2d64        277815483075.dkr.ecr.us-west-2.amazonaws.com/metabiota/rstudio:latest   "/init"             11 seconds ago      Up 10 seconds       0.0.0.0:3838->3838/tcp, 0.0.0.0:8787->8787/tcp   rstudio_test
```

So the container is up and runnign on port 8787 on the local host of the new machine that has been
built. We can retrieve the IP of that machine using the `docker-machine ls` command.

```{sh, eval=FALSE}
docker-machine ls
```

```
NAME         ACTIVE   DRIVER       STATE     URL                        SWARM   DOCKER    ERRORS
awsRStudio   *        amazonec2    Running   tcp://35.161.66.147:2376           v1.12.2   
default      -        virtualbox   Stopped                                      Unknown 
```

This implies that we can navigate to the following address and we should see
the RStudio-Server login page:

http://35.161.66.147:8787/auth-sign-in

Your IP address will be different.

## SSH'ing into container via terminal

Sometimes it is important to be able to SSH into the container via the terminal.
This is easier to do through the `docker exec` command.

```{sh, eval=FALSE}
docker exec -it rstudio_test /bin/bash
```

The flag **-it** represents interactive and through the terminal.

The result is that you are logged into the terminal as `root`.

```
root@f00fa3ff2d64:~# 
```

To exit the terminal, simply type `exit` and you will return to your local session.

## Deleting containers, images, machines

Once you are finished with the machine, you might want to clean up and shut everything down.
This is especially true if you are using an AWS instance and paying by the hour. Not all of
the following steps are necessary but they are shown for reference.

### Delete Container

```{sh, eval=FALSE}
docker kill rstudio_test
docker rm rstudio_test
```

### Delete Image

```{sh, eval=FALSE}
docker rmi 277815483075.dkr.ecr.us-west-2.amazonaws.com/metabiota/rstudio:latest

or

docker rmi rocker/rstudio::latest
```

### Delete Machine

```{sh, eval=FALSE}
docker-machine rm -f awsRStudio
```

And just to double check, you can login into the AWS account to confirm that the instance
has been deleted.

https://us-west-2.console.aws.amazon.com
